Quick Research Review

### 2018-11-05 presentations/Su_DL_110618.tex
stash Daniel Visit ppt

    \documentclass[11pt,
        usenames, % allows access to some tikz colors
        dvipsnames % more colors: https://en.wikibooks.org/wiki/LaTeX/Colors
    ]{report}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[margin=1in]{geometry} % page geometry
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\;\upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

    \newtheorem{theorem}{Theorem}[section]

@@

\def\Snospace~{\S{}} % hack to remove the space left after autorefs
\renewcommand*{\sectionautorefname}{\Snospace}
\renewcommand*{\appendixautorefname}{\Snospace}
\renewcommand*{\figureautorefname}{Fig.}
\renewcommand*{\equationautorefname}{Eq.}
\renewcommand*{\tableautorefname}{Tab.}

\onehalfspacing

\pagestyle{fancy}
\rfoot{Yubo Su}
\cfoot{\thepage/\pageref{LastPage}}

\title{Research Notes}
\author{Yubo Su}
\rhead{}

@@

\tableofcontents

\newpage

\chapter{Preliminary Problems}

To get an intuition for how Dedalus and fluid mechanics works, we will solve
some toy problems. Recall fluid equations in the presence of a uniform
gravitational field $\vec{g} = -g\hat{z}$:
\begin{equation}
    \begin{split}
        \pd{\rho}{t} + \vec{\nabla} \cdot \p*{\rho \vec{u}} &= 0,\\
        \rd{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} - \vec{g} &= 0.
    \end{split}\label{eq:fluid_eq}
\end{equation}

In the incompressible limit, $\rd{\rho}{t} = 0$, which implies $\vec{\nabla}
\cdot \vec{u} = 0$. We use subscripts to indicate perturbed quantities, $Q_0$ is
background and $Q_1$ is perturbed. We will generally use $\vec{u}_0 = 0$ unless
otherwise noted. We will also generally assume symmetry along all axes except
$z$ the vertical axis.

In the incompressible limit, the fluid equations become
\begin{equation}
    \begin{split}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\rho_1}{t} + u_{1z}\pd{\rho_0}{z} &= 0,\\
        \pd{\vec{u}_1}{t} + \frac{1}{\rho_0}\vec{\nabla}P_1
            + \frac{\rho_1 g \hat{z}}{\rho_0} &= 0.
    \end{split}\label{eq:lin.incomp}
\end{equation}
We have used $\vec{\nabla}P_0 = -\rho_0 g \hat{z}$ in the absence of
perturbations.

\section{Incompressible, No Gravity}

We note that in the no gravity limit that $\rho_1$ does not have an effect on
other dynamical variables, so the equations of motion we must solve are
\begin{equation}
    \begin{split}
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P_1}{\rho_0} &= 0.
    \end{split}\label{eq:lin.no_g_eom}
\end{equation}
We can take the divergence of the momentum equation and substitute the
continuity equation to get $\nabla^2 P = 0$.

\subsection{Dirichlet BCs}

This is a Laplace equation, which we've solved countless times. Imposing
periodic boundary conditions in the $x$ direction and $P_1(z = L) = 0, P_1(z =
0) = \mathcal{P}(x, t)$, we obtain eigenfunctions
\begin{equation}
    \begin{split}
        P_{1, n}(x, z, t) &=
            \frac{\mathcal{P}_n(t)}{\sinh(k_nL)}e^{ik_nx}\sinh\p*{k_n(L - z)},\\
        u_{1x, n}(x, z, t) &=
            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{x}\;\mathrm{d}t,\\
        u_{1z, n}(x, z, t) &=
            \int\limits^t -\frac{1}{\rho_0}\pd{P_{1, n}}{z}\;\mathrm{d}t.
    \end{split}\label{eq:incomp_nog}
\end{equation}
We define $k_n = \frac{2\pi n}{L}, n \geq 0$ and $\mathcal{P}(x, t) =
\sum\limits_n \mathcal{P}_n(t)e^{ik_nx}$.

Thus, if we impose BCs $\mathcal{P}(x, t) = \sin \frac{2\pi x}{L}$ and start
with initial conditions such that all quantities are zero, we would expect after
transients die out that
\begin{equation}
    \begin{split}
        P(x, z, t) &= \frac{\sin \frac{2\pi x}{L}}{\sinh 2\pi}
            \sinh \p*{2\pi \frac{L - z}{L}},\\
        u_{1x}(x, z, t) &= -\frac{2\pi t}{L\rho_0}
            \frac{\cos \frac{2\pi x}{L}}{\sinh 2\pi}\sinh \p*{2\pi \frac{L -
                z}{L}},\\
        u_{1z}(x, z, t) &= +\frac{2\pi t}{L\rho_0}
            \frac{\sin \frac{2\pi z}{L}}{\sinh 2\pi}\cosh \p*{2\pi \frac{L -
                z}{L}}.
    \end{split}\label{eq:incomp_nog_sol}
\end{equation}
This is in good agreement with the results, presented in \autoref{fig:no_g}.
Note that $P$ is constant while $\vec{u}$ increases linearly in time, and we
observe the expected $\sim \sin x \sinh \frac{L - z}{z}$ dependence. In fact,
$u_{1x}, u_{1z}$ are exactly $\frac{2\pi}{10}$ at $t = 1$.
\begin{figure}[!h]
    \centering
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t50.png}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t50.png}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t50.png}
    \end{subfigure}

    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_P_t100.png}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_ux_t100.png}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../sims_old/2d_0_no_g/no_g_uz_t100.png}
    \end{subfigure}
    \caption{$P, u_x, u_z$ at $t = 0.5$ and $t = 1$ for $\rho_0 = 1$. We choose
    a $L = 10$ square domain.}\label{fig:no_g}
\end{figure}

\subsection{Things to Note}

It is worth noting that, since our \autoref{eq:lin.no_g_eom} reduced to a Laplace
equation, we needed two $z$ BCs and two $x$ BCs (periodic BCs amount to equating
the value and derivative of the function). This is in agreement with the
observation that the original \autoref{eq:lin.no_g_eom} had two derivatives in
$x, z$ apiece, so we needed two BCs each.

It is also worth seeing from our solution that $P$ immediately goes to the
equilibrium solution. This is not surprising since in the incompressible limit,
sound speed goes to infinity which is the timescale on which the pressure field
adjusts to net forces. Thus, the dynamics solely arise from the static pressure
field pushing the velocity field to equilibrium.

We chose time-independent $\mathcal{P}(x, t)$, but it is clear that whatever
$\mathcal{P}(x, t)$ we choose, the time dependence propagates to the velocities
by way of an integral. If we had instead chosen to take Fourier transform $t \to
\omega$, we would have had to integrate the boundary condition against the
eigenfunctions for each of the $\omega$, which is still easily computable, to
get the full $u_{1x}(x, z, t)$. We consider this in preparation for when we can
only solve for a set of $\vec{k}, \omega$ in the next problem.

The next thing we would have wanted to do is solve a problem with radiative BCs,
but we need to have wave solutions, which are missing in the absence of gravity.
We thus move on to the next configuration.

\section{Incompressible, Stratified w/ Gravity}

\subsection{Eigenfunctions}

Let's restore the $\rho_1 g$ term now. For funsies, we begin by solving for
arbitrary stratification $\rho_0(z)$ first. The fluid equations to first order
reduce to
\begin{equation}
    \begin{split}
        \pd{\rho_1}{t} + \vec{u}_1 \cdot \p*{\vec{\nabla}\rho_0} &= 0,\\
        \vec{\nabla} \cdot \vec{u}_1 &= 0,\\
        \pd{\vec{u}_1}{t} &= -\frac{\vec{\nabla}P_1}{\rho_0}
            - \frac{\rho_1 g}{\rho_0}
    \end{split}\label{eq:lin_incomp}
\end{equation}
We expect there to be some $z$ dependence in the amplitude, so we substitute
variables of form $e^{i(kx - \omega t)}$ and do not specify the $z$ dependence.
This gives us
\begin{equation}
    \begin{split}
        -i\omega \rho_1 - u_{1z}\pd{\rho_0}{z} &= 0,\\
        iku_{1x} + \pd{u_{1z}}{z} &= 0,\\
        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\\
        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
            \frac{\rho_1 g}{\rho_g} &= 0.
    \end{split}
\end{equation}
We substitute $N^2 = -\frac{g}{\rho_0}\pd{\rho_0}{z}$ to obtain
\begin{subequations}
    \begin{align}
        -i\omega \rho_1 - u_{1z}\frac{\rho_0N^2}{g} &= 0,\label{eq:lin.cont}\\
        iku_{1x} + \pd{u_{1z}}{z} &= 0,\label{eq:lin.div}\\
        -iw u_{1x} + \frac{ik_x P_1}{\rho_0} &= 0,\label{eq:lin.mom_x}\\
        -iw u_{1z} + \frac{1}{\rho_0}\pd{P_1}{z} +
            \frac{\rho_1g}{\rho_0} &= 0.\label{eq:lin.mom_z}
    \end{align}
\end{subequations}
Eliminating $u_{1x}$ by substituting~\eqref{eq:lin.div}
into~\eqref{eq:lin.mom_x} and $\rho_1$ by substituting~\eqref{eq:lin.cont}
into~\eqref{eq:lin.mom_z} give
\begin{subequations}
    \begin{align}
        i\omega \pd{u_{1z}}{z} + \frac{k_x^2 P_1}{\rho_0} &= 0,
            \label{eq:lin.eq1}\\
        \p*{\omega^2 - N^2 }u_{1z} + \frac{i\omega}{\rho_0}\pd{P_1}{z} &= 0.
            \label{eq:lin.eq2}
    \end{align}
\end{subequations}
Finally, we multiply~\eqref{eq:lin.eq1} with $\rho_0$ and differentiate
$\mathrm{d}z$ and combine with~\eqref{eq:lin.eq2} to give
\begin{equation}
    \rtd{u_{1z}}{z} + \frac{1}{\rho_0}\pd{\rho_0}{z}\pd{u_{1z}}{z}
        + k_x^2\p*{\frac{N^2}{\omega^2} - 1}u_{1z} = 0.\label{eq:lin.incomp.gen}
\end{equation}

Let's now pick stratification $\rho \propto e^{-z/H}$
\autoref{eq:lin.incomp.gen} clearly has exponential solutions $e^{\kappa z}$ for
\begin{equation}
    \kappa^2 - \frac{\kappa}{H} + k_x^2\p*{\frac{N^2}{\omega^2} - 1} = 0.
\end{equation}
We permit complex $\kappa = \frac{1}{2H} + ik_z$, and from the above clearly
\begin{align}
    k_z^2 &= -\frac{1}{4H^2} + k_x^2\p*{\frac{N^2}{\omega^2} - 1},\nonumber\\
    \omega^2 &= \frac{N^2k_x^2}{k_x^2 + k_z^2 +
        \frac{1}{4H^2}}.\label{eq:strat_disp}
\end{align}
Thus the eigenfunctions are
\begin{equation}
    \begin{split}
        u_{1z} &= e^{z/2H} e^{i(k_zz + k_xx - \omega t)},\\
        u_{1x} &= -\frac{k_z + i/2H}{k_x} u_{1z},\\
        \rho_1 &= \frac{i \rho_0}{H\omega} u_{1z},\\
        P_1 &= -\frac{\rho_0 \omega}{k_x^2}\p*{k_z + i/2H}u_{1z}.
    \end{split}\label{eq:strat_eigens}
\end{equation}

\subsection{Analytically Solving an IVP, Dirichlet + Driving BCs}

We will analyze everything in terms of $u_{1z}$ since it has the simplest form;
note that when actually choosing the BCs we will have to consider the gauge
freedom of $P$ and some considerations we defer to the computational section.

Currently, we have a set of eigenfunctions
\begin{equation}
    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
        = e^{z/2H} e^{i(k_zz + k_xx - \omega t)}.
        \label{eq:strat.uz_eigen}
\end{equation}
Note that $u_{1z}$ is really only a function of two parameters rather than the
three $(k_x, k_z, \omega)$, since the three are related by dispersion relation
\autoref{eq:strat_disp}.

Now, we implement BCs. Consider domain $x, z \in [0, L]$. We will use periodic
BCs again in $x$, so then $k_{x, n} = \frac{2\pi n}{L}, n \geq 0$. Then, we will
require $u_{1z, n}(x, L, t) = 0$, a Dirichlet condition at the top boundary,
which restricts us to eigenfunctions of form
\begin{equation}
    u_{1z, n}\p*{x, z, t \middle| \vec{k}, \omega}
        = e^{z/2H}e^{i(k_xx - \omega t)}\sin\p*{k_z\p*{L - z}}.
        \label{eq:strat.uz_eigen_l-z}
\end{equation}

Finally, we must choose a BC at $z = 0$. We will choose a general function
$u_{1z}(x, 0, t) = F(x, t)$ where we can decompose
\begin{equation}
    F(x, t) = \int\limits \sum\limits_n \mathcal{F}(k_{x, n}, \omega)
        e^{i(k_{x, n}x - \omega t)}\;\mathrm{d}\omega.
\end{equation}
Matching BCs then gives us general solution for $u_{1z}$ given an arbitrary
driving function
\begin{equation}
    u_{1z}\p*{x, z, t \middle| \vec{k}, \omega}
        = \int\limits \sum\limits_n \mathcal{F}\p*{k_{x, n}, \omega}
            \frac{e^{z/2H}e^{i\p*{k_{x, n}x - \omega t}}\sin \p*{k_z(L - z)}}{
            \sin k_zL}\;\mathrm{d}\omega.
            \label{eq:strat.full_sol}
\end{equation}

For ease of computation, let's pick $F(x, t) = \cos \p*{\frac{2 \pi x}{L} -
\omega_0 t}$, so our full expected solution is (note $A + \epsilon =
Ae^{\epsilon/A} + \mathcal{O}(\epsilon^2)$)
\begin{equation}
    \begin{split}
        u_{1z}\p*{x, z, t \middle | \vec{k}, \omega}
            &= e^{z/2H}\frac{\cos\p*{\frac{2\pi x}{L} - \omega_0 t}
                \sin \p*{k_z(L - z)}}{\sin k_zL},\\
        u_{1x}\p*{x, z, t \middle | \vec{k}, \omega}
            &\approx \frac{k_z}{k_x} e^{z/2H}\frac{
                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
                \sin \p*{k_z(L - z)}
            }{\sin k_zL},\\
        \rho_1\p*{x, z, t \middle | \vec{k}, \omega}
            &\approx \frac{\rho_0}{H\omega} e^{z/2H}\frac{
                -\sin\p*{\frac{2\pi x}{L} - \omega_0 t}
                \sin \p*{k_z(L - z)}
            }{\sin k_zL},\\
        P_1\p*{x, z, t \middle | \vec{k}, \omega}
            &\approx -\frac{\rho_0 \omega k_z}{k_x^2} e^{z/2H}\frac{
                \cos\p*{\frac{2\pi x}{L} - \omega_0 t + \frac{1}{2Hk_z}}
                \sin \p*{k_z(L - z)}
            }{\sin k_zL},\\
    \end{split}\label{eq:strat.part_sol}
\end{equation}
where $k_z: \omega(k_x, k_z) = \omega_0$ by the dispersion relation
\autoref{eq:strat_disp}. Note that $H$ contributes both to the overall
exponential profile and to the phase lag of $u_{1x}$.

\subsection{Computationally Solving an IVP, Dirichlet + Driving BCs}

To solve this computationally with the aforementioned BCs, periodic in $x$,
Dirichlet $0$ at $z = L$ and $\cos \p*{\frac{2\pi x}{L} - \omega_0 t}$ at $z =
0$, we must address the gauge freedom in $P$. This arises because for $k_x = 0$,
the divergence-free condition $\vec{\nabla} \cdot \vec{u}_1 = \pd{u_z}{z} = 0$
specificies $u_{1z}$ up to a constant already, so the bottom BC will fix the
value of $u_z$ when $k_x = 0$.

A different way of phrasing the same argument is as follows. Consider the
discrete $N \times N$ (square for notational simplicity) grid. At the boundary,
there is a list of values $f\p*{\z*{x_i}, z_0}$ that lives in an $N$ dimensional
space. We can thus pick a spanning set of $N$ basis vectors, and for each of
these $N$ vectors, by enforcing $\vec{\nabla} \cdot \vec{u} = 0$ at the boundary
we fix the allowed $f\p*{\z*{x_i}, z_{-1}}$ boundary conditions we can implement.
But there exists a choice of basis vectors for which one of the basis vectors is
constant $f_i\p*{\z*{x_i}, z_0} = C$. For this basis vector, the boundary
condition is fully determined, so we only have $N - 1$ dimensions from which to
choose the BCs $f\p*{\z*{x_i}, z_{-1}}$. Since the dimensionality of a space
cannot depend on the choice of basis vector, the divergence free condition
actually yields an extra degree of freedom.

We can use this extra degreee of freedom to specify $P(z = L) = 0$ so the
oscillations should have zero mean. Then we can just simulate away!

From here we can just tweak the parameters until we get something useful, can
look at the videos. Note that we should simulate until at least $T >
\frac{2L_z}{v_{p, z}}$ the $z$ phase space velocity, so we can capture any
reflections off the boundary. Turns out both Dirichlet and Neumann BCs give
strong reflections that produce standing waves in the $z$ and traveling waves in
the $x$.

\subsection{Phase/Group Velocity}

Let's figure out the analytical forms for the phase, group velocity and the
energy density/power flux.

Let's first consider the phase velocity. We traditionally think about the phase
velocity in 1D $v_{ph} = \frac{\omega(k)}{k}$, but in general it is the function
such that $\vec{k} \cdot \vec{r} - \omega t = \vec{k} \cdot \p*{\vec{r} -
\vec{v}_{ph}t}$ is constant (the phase of the wave). Thus, it must satisfy
$\vec{k} \cdot \vec{v}_{ph} = \omega$, and so one sensible choice is
\begin{equation}
    \vec{v}_{ph} = \frac{\omega \hat{k}}{\abs{\vec{k}}}
        = \frac{\omega \vec{k}}{\abs{\vec{k}}^2}.
        \label{eq:v_ph}
\end{equation}

For our stratified atmosphere problem, $\omega^2 = \frac{N^2 k_x^2}{k^2 +
\frac{1}{4H^2}}$. This corresponds to $\vec{v}_{ph} = \frac{Nk_x}{\sqrt{k_x^2 +
k_z^2 + 1/4H^2}} \vec{k}$.

On the other hand the group velocity is given $\vec{v}_g =
\pd{\omega}{k_i}\hat{\imath}$. In our problem, $v_{g, z} =
-\frac{Nk_xk_z}{\p*{k_x^2 + k_z^2 + 1/4H^2}^{3/2}}$ while $v_{g, x} =
\frac{N}{\sqrt{k_x^2 + k_z^2 + 1/4H^2}} - \frac{Nk_x^2}{\p*{k_x^2 +
k_z^2 + 1/4H^2}^{3/2}} = \frac{N\p*{k_z^2 + 1/4H^2}}{\p*{k_x^2 +
k_z^2 + 1/4H^2}^{3/2}}$.

\subsection{Energy/Power Flux}

To compute the energy and power flux of the wave, we recall that for a general
fluid the energy conservation equation reads
\begin{equation}
    \pd{}{t}\p*{\frac{1}{2}\rho v^2 + \rho \epsilon}
        = \vec{\nabla} \cdot \p*{
            \rho \vec{v}\p*{v^2 + \epsilon + \frac{P}{\rho}}
        },
\end{equation}
where $\epsilon$ is the internal energy. But since $P = (\gamma -
1)\rho\epsilon$ and we take $\gamma \to \infty$ incompressible limit, $\epsilon
= 0$, so the energy of the wave is just $\frac{1}{2}\p*{\rho_0 + \rho_1} u_1^2$
and the power flux is $\p*{\rho_0 + \rho_1}\vec{u}_1u_1^2 + P_1\vec{u}_1$.

Another formula given by Sutherland 2011 is $\pd{\ev*{E}}{t} = v_{g, z}
\ev*{E}$ where we average over $x$ one wavelength. I wasn't able to show
that these two agree in general; Sutherland's formula seems to be derived for a
traveling wavepacket whereas we have no such thing. We do not use this
expression since the Landau \& Lifshitz expression works well.

\section{Supressing Reflection}

Sourced from \url{https://people.maths.ox.ac.uk/trefethen/6all.pdf}. There are
largely two ways to do this. The first is to add a damping zone, a region in
which $\dot{q} = -q/\tau_d(z)$ where $\tau_d(0) = 0$ and increases to some
non-small number for some $z_0$ beyond which we want damping. One form, chosen
in Ryan and Dong's paper, is to use multiplicative factor $f(z) = \max\s*{0, 1 -
\frac{\p*{z - z_b}^2}{\p*{z_d - z_b}^2}}$ where $z_b$ is where one begins
supression and $z_b$ is the boundary. Then a dynamical variable $q$ can be
supressed via $\dot{q} = (\dots) - f(z)\frac{q - q_0}{\tau}$ where $\tau$ is the
dynamical timescale on which to damp and $q_0$ is the value to damp to. We will
apply this only to the velocity variables with an eye to extending this approach
to the nonlinear regime, where the velocity variables should still be damped to
zero but $\rho, P$ must capture the stratification and so cannot easily be
artificially damped (maybe? Revisit if this does poorly compared to damping all
linear variables, and damp the nonlinear variables to their
equilibrium/stratified values).

\subsection{First Order}

In order to compute the reflecting boundary condition, we must find a
sequence of differential operators that well approxmiates the system of PDEs. We
can get this from the dispersion relation; since our radiating boundary is along
$z$, we should try to solve for $k_z$ in the dispersion relation to get a
pseudodifferential operator form for $\pd{}{z}u_z$ (we take $k_xH \ll 1$ as well
to simplify):
\begin{align}
    k_z^2 &= \frac{N^2}{\omega^2}k_x^2 - k_x^2 - \frac{1}{4H^2},\nonumber\\
        &\approx k_x^2\p*{\frac{N^2}{\omega^2} - 1},\\
    \pd{u_z}{z} &\approx -\pd{u_z}{x}\sqrt{\frac{N^2}{\omega^2} - 1}.
\end{align}
We pick the negative sign in accordance with an outgoing wave, per the group
velocity formulae for $k_z > k_x$.

From the numerical simulations, it is clear that reflections are well supressed
initially, but the the simulation blows up. This is because our supression is
imperfect (differential approximation to a pseudodifferential operator) and
since our system is dissipation free, reflection grows. Moreover, since any
reflected components seem to have different $k_x, k_z, \omega$, the second time
they're incident on the boundary, our above condition becomes wildly inaccurate.

\section{The Difficulty of the Incompressible, Nonlinear Problem}

We consider the problem where both advective terms are kept and $\rho_1 \sim
\rho_0$. We write down thus nonlinear fluid equations
\begin{equation}
    \begin{split}
        \pd{\rho}{t} + \vec{u} \cdot \p*{\vec{\nabla}\rho} &= 0,\\
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\vec{u}}{t} &= -\p*{\vec{u} \cdot \vec{\nabla}} \vec{u}
            - \frac{\vec{\nabla}P}{\rho} - g\hat{z}.
    \end{split}\label{eq:nonlin_incomp}
\end{equation}
Note that we cannot even subtract off hydrostatic equilibrium anymore since
$\rho$ can deviate greatly from the normal $\rho_0e^{-z/H}$!

\subsection{Rearrangement for Dedalus}

In order to simulate this at all in a spectral program, recall the way that
nonlinear PDEs are decomposed in a spectral code. Given a phase space $Q$,
consider first a system of PDEs expressible in form
\begin{equation}
    \dot{Q} + \bm{L}Q + f(Q) = g
\end{equation}
where $\dot{Q}$ is the time derivative, $\bm{L}$ is some linear operator, $f(Q)$
are nonlinear terms and $g$ is an arbitrary driving function. Spectral codes
work by re-casting the operators to be purely algebraic in an appropriate
spectral basis consisting of some $\varphi_i$ trial and $\xi_i$ test functions.
The \emph{tau spectral method} that Dedalus uses considers using the trial
functions as the test functions, so the effective decomposition that is
performed is
\begin{equation}
    \bra*{\varphi_i}\dot{Q} + \bm{L}Q \ket*{\varphi_j}
        = \dot{Q}_i + L_{ij}Q_q = \bra*{\varphi_i}-f(Q)  + g \ket*{\varphi_i}.
\end{equation}
It is then clear that $-f(Q) + g$ can just be treated as inhomogeneous source
terms, and over a short $\Delta t$ it can be approximated via $-f(Q_0) + g$.
This first-order inhomegeneous ODE then admits closed form exact solutions
$e^{-L_{ij}t}\p*{\dots}$.

One can imagine, for this matter, that $\pd{}{t}$ can be wrapped in the $L_{ij}$
operator and the timestepping computed implicitly; I believe this is what
Dedalus does, implicitly timestepping the left hand side and explicitly the
inhomogeneous terms.

Both of these procedures require $\bm{L}$ linear operator or $\pd{}{t} + \bm{L}$
to be full rank, or invertible. This is enforced analytically by requiring the
same number of BCs as derivatives, and numerically this is little different.
This is why it is possible to have components of $Q$ that are not differentiated
$\partial_t$ but still have well-defined time evolutions; the only condition
required is that the linear terms of the PDE make up a full-rank operator.
Hence, in our linear incompressible equations, we permitted use of $\vec{\nabla}
\cdot \vec{u}_1 = 0$ in lieu of a $\pd{P}{t}$ explicit equation of motion; so
long as the gauge for $P$ (which only appears in derivatives) is properly set,
the operator is full rank and numerically well-defined.

>>>>

    \documentclass[dvipsnames]{beamer}
    \usetheme{Madrid}
    \usefonttheme{professionalfonts}
    \usepackage{
        amsmath,
        amssymb,
        fouriernc, % fourier font w/ new century book
        fancyhdr, % page styling
        lastpage, % footer fanciness
        hyperref, % various links
        setspace, % line spacing
        amsthm, % newtheorem and proof environment
        mathtools, % \Aboxed for boxing inside aligns, among others
        float, % Allow [H] figure env alignment
        enumerate, % Allow custom enumerate numbering
        graphicx, % allow includegraphics with more filetypes
        wasysym, % \smiley!
        upgreek, % \upmu for \mum macro
        listings, % writing TrueType fonts and including code prettily
        tikz, % drawing things
        booktabs, % \bottomrule instead of hline apparently
        cancel % can cancel things out!
    }
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures

    \newcommand*{\scinot}[2]{#1\times10^{#2}}
    \newcommand*{\dotp}[2]{\left<#1\,\middle|\,#2\right>}
    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\ptd}[2]{\frac{\partial^2 #1}{\partial#2^2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\pvec}[1]{\vec{#1}^{\,\prime}}
    \newcommand*{\svec}[1]{\vec{#1}\;\!}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\ang}[0]{\;\text{\AA}}
    \newcommand*{\mum}[0]{\\;upmu \mathrm{m}}
    \newcommand*{\at}[1]{\left.#1\right|}

@@

\title{Quick Research Review}
\author{Yubo Su}
\date{Nov. 06, 2018}

@@

\begin{frame}
    \frametitle{Objectives}

    \begin{itemize}
        \item Continuous train of IGW excited in stratified WD atmosphere by
            binary companion, grows due to stratification, reaches nonlinear
            amplitudes $\xi_z k_z \gtrsim 1$ before the peak of the envelope.

        \item Seek long-term behavior of nonlinear dissipation, whether
            steady-state or periodic.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Numerical Simulations, No Shear}
    \framesubtitle{Setup}

    \begin{itemize}
        \item 2D, Fourier $\times$ Chebyshev, damping zones, ran at both $128
            \times 512$ and $256 \times 1024$ over $3H \times 12H$ grid, where
            $H$ is scale height.

        \item Target waves $k_x = \frac{2\pi}{3H}, k_z = \frac{2\pi}{H}, \omega
            \approx 0.32N$, $N$ B-V frequency.

        \item Wave is excited by volumetric forcing
            $e^{-z^2/2\sigma^2}\cos\p*{k_xx - \omega t}$ with $\sigma =
            \frac{3H}{32} \lesssim \frac{1}{k_z}$.

        \item Regularized by numerical viscosity $\nu \gtrsim
            \frac{\omega}{k_{\max}k_z}$.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Numerical Simulations, No Shear}
    \framesubtitle{Results}

    \begin{itemize}
        \item High resolution ($\nu = 4\frac{\omega}{k_{\max}k_z}$):
            \lstinline{nl3.mp4}. Note that mean flow starts to build up at the
            outgoing damping zone, and once it hits criticality it starts to
            move down while continuing to steepen. Eventually, it stops moving
            and produces reflection below, while leaving behind a sinusoidal
            mean flow above.

        \item Low resolution ($\nu = 3\frac{\omega}{k_{\max}k_z}$):
            \lstinline{nl3_lowres.mp4}. Qualitatively similar, but the mean flow
            forms later.

        \item Two hypotheses:
            \begin{itemize}
                \item Damping layer is causing mean flow to begin building up
                    like Daniel said; how can we fix? Seems like low $k_z$ modes
                    (which have higher $v_{ph}$).

                \item Further growth/propagation is critical layer absorption?
                    Eventual reflection might build up when shear flow is too
                    steep $\mathrm{Ri} \gtrsim 1/2$ or WKB breaks down
                    $\pd{k_z}{z} \sim \frac{k_z}{\lambda_z}$.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Numerical Simulations, Shear Flow}
    \framesubtitle{Setup}

    \begin{itemize}
        \item 2D, Fourier $\times$ Fourier, damping zones, running at $64 \times
            256$. No height stratification, domain is $ H \times H$ (though $H$
            is now physically meaningless).

        \item Target waves $k_x = \frac{2\pi}{H}, k_z = \frac{20\pi}{H}, \omega
            \approx 0.1N$, $N$ B-V frequency. Volumetric forcing.

        \item Regularized by $\nabla^6$ hyperviscosity.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Numerical Simulations, Shear Flow}
    \framesubtitle{Results}

    \begin{itemize}
        \item Initialize w/ shear flow and let evolve. Try both narrow and wide
            profiles (\lstinline{vstrat*.mp4}).

        \item Conclusion: attenuated transmission when thin, barely-critical
            layer, smooth WKB-like wavelength shortening when broad, critical
            flow, but steep $\pd{U_0}{z}$ gives reflection!

        \item Recall Booker \& Bretherton result
            \begin{align}
                T &\sim \exp\s*{-\pi \sqrt{\frac{N^2}{(U_0')^2} - \frac{1}{4}}}
                & R &\sim
                    \exp\s*{-2\pi \sqrt{\frac{N^2}{(U_0')^2} - \frac{1}{4}}}.
            \end{align}
            WKB criterion $\Rightarrow \frac{k_z^2}{N}U_0' \frac{\lambda^2}{2}
            \ll 1$.

        \item Winters d'Asaro had a local $\mathrm{Ri} \sim 0.5$.
    \end{itemize}
\end{frame}

\end{document}
